### 2023-01-24 document_phases/content/02_model.tex
tex

\chapter{Model}
\label{chap:model}

This work focuses on the extended Hubbard model on a 2D square lattice with periodic boundary conditions.
Its general Hamiltonian in position space is given by

\begin{equation}
    \begin{aligned}
        H_\text{H} =&-t \sum_{\langle r,s \rangle,\sigma} \left( c_{r\sigma}^\dagger c_{s\sigma} + \text{h.c.} \right) 
                        + U \sum_r n_{r\uparrow} n_{r\downarrow} \\
                    &+ \frac{V}{2} \sum_{\langle r,s \rangle} \left( n_{r\uparrow} + n_{r\downarrow} \right) 
                        \left( n_{s\uparrow} + n_{s\downarrow} \right)\,,
    \end{aligned}  
\end{equation}

where $c_{r\sigma}^{(\dagger)}$ represents the standard fermionic annihilation (creation) operator 
and $n_{r\sigma}$ the number operator with spin $\sigma$ on the lattice site $r$.
The constants $t$, $U$, and $V$ govern the strength of the nearest neighbour hopping, 
the onsite interaction, and the nearest neighbour interaction, respectively.
Applying a Fourier transform yields the new momentum space Hamiltonian

\begin{equation}
    \begin{aligned}
        H_k =& \sum_{k \sigma} \epsilon_0 (k) n_{k\sigma} 
                + \frac{U}{N} \sum_{k,p,q} c_{k\uparrow}^\dagger c_{k+q\uparrow} c_{p\downarrow}^\dagger c_{p-q\downarrow} \\
            &+ \frac{V}{2N} \sum_{k,p,q} \sum_{\sigma, \sigma'} \sum_{\delta} \cos(q\delta) 
                    c_{k\sigma}^\dagger c_{k+q\sigma} c_{p\sigma'}^\dagger c_{p-q\sigma'}\,.
    \end{aligned} 
\end{equation}

Here, $k$, $p$, and $q$ are momenta, $\delta$ references the vectors connecting nearest neighbours within the lattice,
and $N$ is the total number of lattice sites.
The diagonal part is given by

\begin{equation}
    \epsilon_0 (k) := -2t \sum_\delta \cos(\delta k)\,.
\end{equation}